Self-assembly of colloid-cholesteric composites: A route to switchable optical materials

### 2013-10-03 papers/2012-sandwich/CholColloids.tex
updated text
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig/trunk@2173 40876441-3496-0410-a613-a2fca35b2ae2

When spherical colloidal particles are mixed into a nematic liquid crystal, 
they disrupt the orientational order in the fluid and create defects (disclination lines) in the nematic close to their surfaces. To minimize the free energy cost, it is generally advantageous for defects to be shared:
therefore the colloidal particles have a generic tendency to aggregate. Such aggregates may further 
self-assemble into lines~\cite{wiresmiha}, 2D crystals~\cite{zumer}, 
planar structures~\cite{tanaka}, or 3D amorphous glasses~\cite{tiffany}.
Besides being of fundamental interest to materials science, these
structures have tunable elastic and optical properties~\cite{stark}. Hence 
they offer exciting prospects for applications as biosensors~\cite{abbott}, or
as new-generation devices~\cite{colloiddevice,electrophoresis}.

@@

A second compelling reason to study colloid-BP composites is that, even without particles, BPs can show several competing metastable free energy minima,
corresponding to different topologies of the defect network~\cite{adriano}, with a strong dependence on applied electric or magnetic fields~\cite{henrichfield}.
% 
Adding particles is likely to further enrich the free energy landscape, creating composites that might be promising candidates for bistable or multistable devices, in which energy is needed only to switch optical properties and not to maintain them. (This is the e-paper paradigm~\cite{epaper}, and can lead to huge energy savings.)
To explore such metastability, and switching strategies between states, we require multi-unit-cell, time-dependent simulations. These go beyond previous studies of colloid-BP composites~\cite{miha} but follow comparable simulations of pure BPs~\cite{bp3,henrichfield,domaingrowth}, and isolated and dimeric colloids in cholesterics~\cite{juho1,juho2}. Our methodology and simulation parameters are summarized in the Supporting Information.

@@

For a single colloidal particle in a nematic fluid, $w$ controls the formation of a local topological defect at the particle surface~\cite{stark}, with weak distortion of the nearby fluid at small $w$. In the colloid-BP system, $w$ also plays a determining role in the final composite structure. When $w$ is small (Fig.~ 1A,B), the attraction of particles to the disclination lattice causes
 the lattice to become covered by particles, with little disruption of its
 long-range order. Thus a templated colloidal crystal is created, as first reported in~\cite{miha}. Interestingly, the weak but finite anchoring apparently creates an elastic force between the colloids which leads to very slow dynamics (Fig. S4) and the formation of small colloidal lines within disclinations, especially for dilute samples (Fig. 1A). It seems likely that polymer-stabilised  blue phases~\cite{kikuchi} work by a similar principle to this small $w$ case, with a weak segregation of monomers towards the defect regions.

For large $w$ (Fig.~1C,D), the picture changes dramatically.
Each colloidal particle now disrupts significantly the order of the nearby fluid, with defect formation close to its surface~\cite{stark}. The resulting strong disruption of local liquid crystalline order around each particle leads now to 
a strong interparticle attraction. This restructures the disclination
network completely, destroying the long range order of
the original BPI topology. Most notably, particle aggregation favours
the formation of defect junctions, where four disclinations meet. This structural motif is present in the unit cell of BPII rather than BPI (and also seen in the amorphous BPIII~\cite{bp3}). The particle clusters are disjoint at low volume fractions ($\phi\simeq 1\%$) but they interact elastically via the connecting disclinations. At larger $\phi$, the aggregates join up to form a percolating colloidal cluster which essentially templates the disclinations rather than vice versa.

@@

An interesting question, relevant for applications to devices, is whether 
these structures can be modified, or switched, with electric fields.
Fig.~3 addresses this question, starting from amorphous Blue Phase III,
this time at a higher chirality. Dispersing colloids in this networks
results in the particle occupying the available nodes, where the order
parameter is low (Fig.~3A). Applying a high enough field along the $z$ 
direction leads to a switching to a new phase, where colloidal arrays fit 
within a network with distorted hexagonal cells (Fig.~3B). Remarkably, upon 
field removal this honeycomb structure does not find its way back to the 
configuration before switching on the field, but gets stuck into a metastable 
structure (Fig. S7), with residual anisotropy along the field switching 
direction.
The resulting multistability is indeed a prerequisite for new generation 
switchable electrooptic devices for use in energy-saving electronics-free
appliances, smart glass, and e-paper to cite just a few examples.

@@

\bibitem{electrophoresis}  O. D. Lavrentovich,	I. Lazo, O. P. Pishnyak.
Nonlinear electrophoresis of dielectric and metal spheres in a nematic liquid 
crystal. {\it Nature} {\bf 467}, 947-950 (2010).

@@

@@

\begin{figure}
\centerline{\includegraphics[width=\textwidth]{text-fig3.jpg}}
\caption{Snapshots of the steady states obtained when a dispersion of colloids is inserted into a thermodynamically stable bulk-BPIII (A), and finally subjected to an electric field along the vertical direction (B). The electric field leads to the formation of a honeycomb structure with hexagonal ordering perpendicular to the field direction. 
Structures correspond to $w=0.23$ and $\phi=1\%$ (see Supporting Information
for full list of parameters). }
\end{figure}

>>>>

When spherical colloidal particles are mixed into a nematic liquid crystal, 
they disrupt the orientational order in the fluid and create defects (disclination lines) in the nematic close to their surfaces. To minimize the free energy cost, it is generally advantageous for defects to be shared:
therefore the colloidal particles have a generic tendency to aggregate. Such aggregates may further 
self-assemble into lines~\cite{wiresmiha}, 2D crystals~\cite{zumer}, 
planar structures~\cite{tanaka}, or 3D amorphous glasses~\cite{tiffany}.
Besides being of fundamental interest to materials science, these
structures have tunable elastic and optical properties~\cite{stark}. Hence 
they offer exciting prospects for applications as biosensors~\cite{abbott}, or
as new-generation devices~\cite{colloiddevice,tanakanatmat}.

@@

A second compelling reason to study colloid-BP composites is that, even without particles, BPs can show several competing metastable free energy minima,
corresponding to different topologies of the defect network~\cite{adriano,fukuda}, with a strong dependence on applied electric or magnetic fields~\cite{henrichfield}.
% 
Adding particles is likely to further enrich the free energy landscape, creating composites that might be promising candidates for bistable or multistable devices, in which energy is needed only to switch optical properties and not to maintain them. (This is the e-paper paradigm~\cite{epaper}, and can lead to huge energy savings.)
To explore such metastability, and switching strategies between states, we require multi-unit-cell, time-dependent simulations. These go beyond previous studies of colloid-BP composites~\cite{miha} but follow comparable simulations of pure BPs~\cite{bp3,henrichfield,domaingrowth}, and isolated and dimeric colloids in cholesterics~\cite{juho1,juho2}. Our methodology and simulation parameters are summarized in the Supporting Information.

@@

For a single colloidal particle in a nematic fluid, $w$ controls the formation of a local topological defect at the particle surface~\cite{stark}, with weak distortion of the nearby fluid at small $w$. In the colloid-BP system, $w$ also plays a determining role in the final composite structure. When $w$ is small (Fig.~ 1A,B; Suppl. Movie 1,2), the attraction of particles to the disclination lattice causes
 the lattice to become covered by particles, with little disruption of its
 long-range order. Thus a templated colloidal crystal is created, as first reported in~\cite{miha}. Interestingly, the weak but finite anchoring apparently creates an elastic force between the colloids which leads to very slow dynamics (Fig. S4) and the formation of small colloidal lines within disclinations, especially for dilute samples (Fig. 1A). It seems likely that polymer-stabilised  blue phases~\cite{kikuchi} work by a similar principle to this small $w$ case, with a weak segregation of monomers towards the defect regions.

For large $w$ (Fig.~1C,D; Suppl. Movie 3,4), the picture changes dramatically.
Each colloidal particle now disrupts significantly the order of the nearby fluid, with defect formation close to its surface~\cite{stark}. The resulting strong disruption of local liquid crystalline order around each particle leads now to 
a strong interparticle attraction. This restructures the disclination
network completely, destroying the long range order of
the original BPI topology. Most notably, particle aggregation favours
the formation of defect junctions, where four disclinations meet. This structural motif is present in the unit cell of BPII rather than BPI (and also seen in the amorphous BPIII~\cite{bp3}). The particle clusters are disjoint at low volume fractions ($\phi\simeq 1\%$) but they interact elastically via the connecting disclinations. At larger $\phi$, the aggregates join up to form a percolating colloidal cluster which essentially templates the disclinations rather than vice versa.

@@

An interesting question, relevant for applications to devices, is whether 
these structures can be modified, or switched, with electric fields.
Fig.~3 addresses this question, starting from amorphous Blue Phase III,
this time at a higher chirality. Dispersing colloids in this networks
results in the particle occupying the available nodes, where the order
parameter is low (Fig.~3A; Suppl. Movie 5). Applying a high enough field along the $z$ 
direction leads to a switching to a new phase, where colloidal arrays fit 
within a network with distorted hexagonal cells (Fig.~3B; Suppl. Movie 6). Remarkably, upon 
field removal this honeycomb structure does not find its way back to the 
configuration before switching on the field, but gets stuck into a metastable 
structure (Fig. S7), with residual anisotropy along the field switching 
direction.
The resulting multistability is indeed a prerequisite for new generation 
switchable electrooptic devices for use in energy-saving electronics-free
appliances, smart glass, and e-paper to cite just a few examples.

@@

\bibitem{tanakanatmat} T. Araki, M. Buscaglia, T. Bellini, H. Tanaka.
Memory and topological frustration in nematic liquid crystals confined in porous materials. {\it Nat. Mat.} {\bf 10}, 303-309 (2011).
%\bibitem{electrophoresis}  O. D. Lavrentovich,	I. Lazo, O. P. Pishnyak.
%Nonlinear electrophoresis of dielectric and metal spheres in a nematic liquid 
%crystal. {\it Nature} {\bf 467}, 947-950 (2010).

@@

\bibitem{fukuda} J.-I. Fukuda, S. Zumer.
Quasi-two-dimensional Skyrmion lattices in a chiral nematic liquid crystal. 
{\it Nat. Comm.} {\bf 2}, 246 (2011).

@@

\begin{figure}
\centerline{\includegraphics[width=\textwidth]{text-fig3.jpg}}
\caption{Snapshots of the steady states obtained when a dispersion of colloids is inserted into a thermodynamically stable bulk-BPIII (A), and finally subjected to an electric field along the vertical direction (B). The electric field leads to the formation of a honeycomb structure with hexagonal ordering perpendicular to the field direction. 
Structures correspond to $w=0.23$ and $\phi=5\%$ (see Supporting Information
for full list of parameters). }
\end{figure}
